---
title: "R Notebook"
output: html_notebook
---

```{r}
install.packages("lmtest")
library(lmtest)
```

```{r}
diabetes <-read_csv("/home/participant/R/data/diabetes.csv")
```

#Get the glm for the diabetes data again
```{r}
glm_dia <- glm(test_result ~ glucose * diastolic,
                  family = "binomial",
                  data = diabetes)
```

#Look at the model
```{r}
summary(glm_dia)
```
#Get the nul glm for the diabetes data
```{r}
glm_null <- glm(test_result ~ 1, 
                family = binomial, 
                data = diabetes)
```

#Test between the glm for the doabetes data and the null
```{r}
lrtest(glm_dia, glm_null)
```

#Investigate if the interaction between glucose and diastolic gives an impproved model, first get a model
```{r}
glm_dia_add <- glm(test_result ~ glucose + diastolic,
                  family = "binomial",
                  data = diabetes)
```

#Compare against the original model, it does not improve things
```{r}
lrtest(glm_dia, glm_dia_add)
```

#ANOVA can be used to compare all factors in the model, glucose is the best
```{r}
anova(glm_dia, test="Chisq")
```

#Use a goodness of fit chi squared test to test the model
#Model deviance and model degrees of freedom (df) are needed
```{r}
1 - pchisq(glm_dia$deviance, glm_dia$df.residual)
```

#AIC values can tell us about overall model quality
#They are calculated automatically and included in the model summary
```{r}
glm_dia$aic
```

#The R step function can also be used to check the reliability of our model
```{r}
step(glm_dia, test = "LRT")
```

