---
title: "R Notebook"
output: html_notebook
---

```{r}
# install pwr package if needed
install.packages("pwr")

# load the pwr package
library(pwr)
```

#Use to get the effect values needed for experiments from Cohen
```{r}
cohen.ES(test = "t", size = "medium")
```

#Carry out power test for two sample t test
```{r}
pwr.t.test(n = 30, sig.level = 0.05, power = 0.8,
           type = "two.sample", alternative = "two.sided")
```
#Use fish data again and have a look
```{r}
# read in the data
fishlength <- read_csv("data/CS1-twosample.csv")

# visualise the data
fishlength %>% 
  ggplot(aes(x = river, y = length)) +
  geom_boxplot()
```
#Perform t test
```{r}
# perform t-test
t.test(length ~ river,
       data = fishlength,
       var.equal = TRUE)
```

Carry out cohens_d function on fish data to get effect size
```{r}
cohens_d(length ~ river,
         var.equal = TRUE,
         data = fishlength)
```

Carry out power analysis on fish data using effect size
```{r}
pwr.t.test(d = 0.94, power = 0.8, sig.level = 0.05,
           type = "two.sample", alternative = "two.sided")
```
#We only need 19 fish!!

#Read in and look at lobster data again
```{r}
# read in the data
lobsters <- read_csv("data/CS2-lobsters.csv")
# visualise the data
ggplot(lobsters,
       aes(x = diet, y = weight)) +
  geom_boxplot()
```

##Carry out linear model and use for ANOVA of fish data
```{r}
# define the linear model
lm_lobster <- lm(weight ~ diet,
                 data = lobsters)

# perform ANOVA on model
anova(lm_lobster)
```

#Use Cohen's f2 to get effect size from R2 value from linear model
```{r}
# get the effect size for ANOVA
R2 <- summary(lm_lobster) %>%
    glance() %>% 
    pull(r.squared)

# calculate Cohen's f2
R2 / (1 - R2)
```

##Carry out power calculation. U is the numerator degress of freedom (one less than number of parameters, i.e 3 fish -1 = 2)
```{r}
pwr.f2.test(u = 2, f2 = 0.219,
            sig.level = 0.05, power = 0.8)
```

#Number of observations needed = U(numerator df) + V(denominator df) + 1

```{r}
# read in the data
fish <- read_csv("data/CS1-onesample.csv")
# visualise the data
ggplot(fish,
       aes(x = river, y = length)) +
  geom_boxplot()
```

```{r}
t.test(fish$length,
       mu = 20, alternative = "two.sided")
```

```{r}
cohens_d(length ~ 1,
         mu = 20,
         data = fish)
```

```{r}
pwr.t.test(d = -0.659, power = 0.9, sig.level = 0.01,
           type = "one.sample")
```

```{r}
# read in the data
cortisol <- read_csv("data/CS1-twopaired.csv")
# visualise the data
```

```{4}
ggplot(fish,
       aes(x = river, y = length)) +
  geom_boxplot()
```

```{r}
cortisol_diff <- cortisol %>%
  pivot_wider(id_cols = patient_id,
              names_from = time,
              values_from = cortisol) %>% 
  mutate(cortisol_change = evening - morning)

cortisol_diff
```


```{r}
t.test(cortisol_diff$morning, cortisol_diff$evening,
       alternative = "two.sided",
       paired = TRUE)
```

```{r}
cohens_d(cortisol ~ time,
       paired = TRUE,
         data = cortisol)
```

```{r}
pwr.t.test(d = -1.159019, power = 0.8, sig.level = 0.01,
           type = "paired")
```

```{r}
# read in the data
shell <- read_csv("data/CS6-shelllength.csv")
# visualise the data
```

```{r}
ggplot(shell,
       aes(x = location, y = length)) +
  geom_boxplot()
```
```{r}
# define the linear model
lm_shell <- lm(length ~ location,
                 data = shell)

# perform ANOVA on model
anova(lm_shell)
```

```{r}
# get the effect size for ANOVA
R2 <- summary(lm_shell) %>%
    glance() %>% 
    pull(r.squared)

# calculate Cohen's f2
effect_size_shell <- R2 / (1 - R2)
```


```{r}
pwr.f2.test(u = 4, f2 = effect_size_shell,
            sig.level = 0.05, power = 0.8)
```

```{r}
epilepsy <- read_csv("data/CS6-epilepsy1.csv")
```

```{r}
ggplot(epilepsy,
       aes(x = age, y = seizure)) +
    geom_point()
```

```{r}
epilepsylm <- lm(seizure ~ age,
                   data = epilepsy)
```

```{r}
R2_epilepsy <- epilepsylm %>%
    glance() %>% 
    pull(r.squared)
```

```{r}
co_f2_epilepsy <- R2_epilepsy / (1 - R2_epilepsy)
```



```{r}
pwr.f2.test(u = 1, f2 = co_f2_epilepsy,
            sig.level = 0.05, power = 0.9)
```

