---
title: "Generalised_Linear_Models"
output: html_document
date: "2024-06-14"
---

```{r setup, include=FALSE}
library(broom)
library(tidyverse)
library(ggResidpanel)
```

#Load in data, check pointed beak, this is categorical either it is or isn't
#This is represented by 0 or 1.
```{r}
early_finches <- read_csv("/home/participant/R/data/finches_early.csv")
head(early_finches$pointed_beak)
```

#Box plot of pointed beak data
```{r}
ggplot(early_finches,
       aes(x = factor(pointed_beak),
          y = blength)) +
  geom_boxplot()
```


#Line of best fit does not work in this case, we have values less or more than 1
```{r}
ggplot(early_finches,
       aes(x = blength, y = pointed_beak)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE)
```

#Test of assumptions is terrible as it is a categorical response
```{r}
lm_bks <- lm(pointed_beak ~ blength,
             data = early_finches)

resid_panel(lm_bks,
            plots = c("resid", "qq", "ls", "cookd"),
            smoother = TRUE)
```
#Insteasd we fit a generalised linear function, this uses a logistic link function that will esnure the reponse variable is between 0 and 1
#Not we have to specify the family of the error distribution to use, here it is binomial
```{r}
glm_bks <- glm(pointed_beak ~ blength,
               family = binomial,
               data = early_finches)
```

```{r}
glm_bks %>% 
  augment(type.predict = "response") %>% 
  ggplot() +
  geom_point(aes(x = blength, y = pointed_beak)) +
  geom_line(aes(x = blength, y = .fitted),
            linetype = "dashed",
            colour = "blue") +
  geom_point(aes(x = blength, y = .fitted),
             colour = "blue", alpha = 0.5) +
  labs(x = "beak length (mm)",
       y = "Probability")
```

```{r}
resid_panel(glm_bks, plots = "cookd")
```

#Check how well our model predicts our data
#This uses the pchisq function and takes the residual deviance and degrees of freedom from the ouput of the GLM object.
```{r}
pchisq(9.1879, 59, lower.tail = FALSE)
```

```{r}
anova(glm_bks, test = "Chisq")
```

#Use the ANOVA function to see if our predictors are significant, this is shown in the Pr(>Chi) column
```{r}
anova(glm_bks, test = "Chisq")
```

#Diabetes data
```{r}
diabetes <-read_csv("/home/participant/R/data/diabetes.csv")
```

```{r}
ggplot(diabetes,
       aes(x = factor(test_result),
           y = glucose)) +
  geom_boxplot()
```


#Box plot of pointed beak data
```{r}
ggplot(diabetes,
       aes(x = glucose, y = test_result)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE)
```

```{r}
glm_diabetes <- glm(test_result ~ glucose,
               family = binomial,
               data = diabetes)
```

```{r}
anova(glm_diabetes, test = "Chisq")
```

```{r}
glm_diabetes %>% 
  augment(type.predict = "response") %>% 
  ggplot() +
  geom_point(aes(x = glucose, y = test_result)) +
  geom_line(aes(x = glucose, y = .fitted),
            linetype = "dashed",
            colour = "blue") +
  geom_point(aes(x = glucose, y = .fitted),
             colour = "blue", alpha = 0.5) +
  labs(x = "Glucose (mM)",
       y = "Probability")
```

```{r}
exp(-5.61173 + 0.03951  * 150) / (1 + exp(-5.61173 + 0.03951  * 150))
```

