---
title: "Nested Random Effects"
output: html_document
date: "2024-07-16"
---

```{r}
library(tidyverse)
library(lme4)
library(broom)
library(broom.mixed)
library(patchwork)
```


```{r}
health <- read_csv("/home/participant/Documents/health.csv")
```

#Create a mixed effect linea model with intercepts only (1|doctor) + (1|hospital)
```{r}
lme_health_intercepts <- lmer(outcome ~ treatment + (1|doctor) + (1|hospital),
                   data = health)

summary(lme_health_intercepts)
```

#Now add random intercepts and slopes for both, this is because the doctor gives both treatments (Look at table to see this)
```{r}
lme_health_slopes <- lmer(outcome ~ treatment + (1 + treatment|doctor) + 
                      (1 + treatment|hospital), data = health)

summary(lme_health_slopes)
```
#New data set where doctors only give one type of treatment
```{r}
health2 <- read_csv("/home/participant/Documents/health2.csv")
```

#In this case we fit only random intercepts for doctor, but both intercepts and slopes for the hospital as each hospital gives both treatments
```{r}
lme_health2_slopes <- lmer(outcome ~ treatment + (1|doctor) + 
                      (1 + treatment|hospital), data = health2)

summary(lme_health2_slopes)
```
#The data above was implicitly encoded (The best way to do it), with a separate ID for each patient
#The pastes dataset is explicitly encoded, we have 30 casks, but no unique cask id.
```{r}
data("Pastes")

head(Pastes)
```

#Just fitting as before confuses R, as R thinks we only have 3 casks; A, B and C.
```{r}
lme_paste <- lmer(strength ~ 1 + (1|batch) + (1|cask), data = Pastes)

summary(lme_paste)
```

#To get around the problem we can add a new column that gives each cask a unique ID
```{r}
Pastes <- Pastes %>% mutate(unique_cask = batch:cask)

head(Pastes)
```

#We can then use the unique ID to get a random intercept
```{r}
lme_paste_implicit <- lmer(strength ~ 1 + (1|batch) + (1|unique_cask),
                  data = Pastes)

summary(lme_paste_implicit)
```

#If we don't want to create a new column, we can just analyse the data explicitly, to do this we tell R how the casks relate to the batch (1|batch:cask)
```{r}
lme_paste_explicit <- lmer(strength ~ 1 + (1|batch) + (1|batch:cask), data = Pastes)

summary(lme_paste_explicit)
```

#This is the same as above but with different sytax, we use (1|batch/cask) here instead
```{r}
lme_paste_shorthand <- lmer(strength ~ 1 + (1|batch/cask), data = Pastes)

summary(lme_paste_shorthand)
```

```{r}
cake <- data("cake")
```


