---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
library(broom) #Needed for augment()
library(rstatix)
library(ggResidpanel)
```


```{r}
pm2_5 <- read_csv("/home/participant/Course_Materials/data/CS5-pm2_5.csv")
```

```{r}
ggplot(pm2_5,
       aes(x = location, y = pm2_5)) +
    geom_boxplot() +
    geom_jitter(width = 0.1, alpha = 0.3)
```

```{r}
pm2_5 %>% 
    select_if(is.numeric) %>% 
    pairs(lower.panel = NULL)
```

```{r}
ggplot(pm2_5,
       aes(x = wind_m_s, y = pm2_5,
           colour = location)) +
    geom_point()
```

```{r}
pm2_5 %>% 
    select_if(is.numeric) %>% 
    cor()
```

##Get linear model
```{r}
lm_pm2_5_full <- lm(pm2_5 ~ avg_temp + location +
                            rain_mm + wind_m_s +
                            wind_m_s:location,
                    data = pm2_5)
```

#Check out effects on the response with ANOVA
```{r}
anova(lm_pm2_5_full)
```
#Rain does not effect pm2_5 so........

##Get a reduced model without rain
```{r}
lm_pm2_5_red <- lm(pm2_5 ~ avg_temp + location + wind_m_s + location:wind_m_s, data = pm2_5)
```

#Visualise the interactions
```{r}
lm_pm2_5_red %>% 
  augment() %>% #Convert to format ggplot can use
  ggplot(aes(x = wind_m_s,
             y = pm2_5, colour = location)) +
  geom_point() +
  geom_smooth(aes(y = .fitted))
```
#Get an addative model, without the interaction of wind and location (we would normally not do this because the interaction term is significant!)
```{r}
lm_pm2_5_add <- lm(pm2_5 ~ avg_temp + location + wind_m_s,
                   data = pm2_5)
```

```{r}
lm_pm2_5_add %>% 
    augment() %>%
    ggplot(aes(x = wind_m_s, y = pm2_5,
               colour = location)) +
    geom_point() +
    geom_smooth(aes(y = .fitted))
```
#Notice the lines are parallel now

##Lets just look at wind
```{r}
lm_pm2_5_wind <- lm(pm2_5 ~ wind_m_s,
                   data = pm2_5)
```

#Visualise using the lm object and reformat using augment
```{r}
lm_pm2_5_wind %>% 
    augment() %>% 
    ggplot(aes(x = wind_m_s, y = pm2_5)) +
    geom_point() +
    geom_smooth(aes(y = .fitted))
```
#Just visualise using geom smooth and carry out the linear model analysis on the fly
```{r}
ggplot(pm2_5, aes(x = wind_m_s, y = pm2_5)) +
    geom_point() +
    geom_smooth(method = "lm", se = FALSE)
```

#Extract co-efficients
```{r}
lm_pm2_5_wind
```

#Lets look at location, it is categorical so if therefore just an ANOVA
```{r}
lm_pm2_5_loc <- lm(pm2_5 ~ location,
                   data = pm2_5)
```

```{r}
lm_pm2_5_loc %>% 
    augment() %>% 
    ggplot(aes(x = location, y = pm2_5)) +
    geom_jitter(alpha = 0.3, width = 0.1) +
    geom_point(aes(y = .fitted), colour = "blue", size = 3)
```
#Check mean are the smae as what the model gives above (The means are the model!)
```{r}
pm2_5 %>% 
    group_by(location) %>% 
    summarise(mean_pm2_5 = mean(pm2_5))
```

#Get the coefficients
```{r}
lm_pm2_5_loc
```

###Look at tree data
```{r}
tree <- read_csv("/home/participant/Course_Materials/data/CS5-trees.csv")
```

```{r}
lm_tree1 <- lm(volume ~ height + girth + height:girth,
                   data = tree)
```

```{r}
lm_tree2 <- lm(volume ~ height + girth,
                   data = tree)
```

```{r}
lm_tree3 <- lm(volume ~ height,
                   data = tree)
```

```{r}
lm_tree4 <- lm(volume ~ girth,
                   data = tree)
```

```{r}
resid_panel(lm_tree1,
            plots = c("resid", "qq", "ls", "cookd"),
            smoother = TRUE)
```

```{r}
resid_panel(lm_tree2,
            plots = c("resid", "qq", "ls", "cookd"),
            smoother = TRUE)
```

```{r}
resid_panel(lm_tree3,
            plots = c("resid", "qq", "ls", "cookd"),
            smoother = TRUE)
```

```{r}
resid_panel(lm_tree4,
            plots = c("resid", "qq", "ls", "cookd"),
            smoother = TRUE)
```





